Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
going home

% IEAs
Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
whose fitness evaluations are performed by humans through an interactive 
system \cite{eiben2015interactive}.
They are usually applied in problems where the fitness function is not known or simply
does not exist, and the result of optimization should fit a certain human need or desire. 
That is why their use-case includes the evolution of objects with subjective characteristics
such visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
or creativity \cite{yu2011cooks}.
% Human Based Computation
These IEC systems are an interesting venue of research, since they have demonstrated 
their ability for effectively 
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
as well as other artifacts in many other domains \cite{ie1}. In those cases when 
human interaction is responsible of other 
aspects of the evolutionary process, some authors classify these IEC methods 
as human-based evolutionary computation \cite{kosorukoff2001human} 
or as human-based computation \cite{quinn2011human}.
% Problems
However, the necessary intervention of humans, brings certain challenges 
to designers of IEC methods namely, human evaluations are slow and expensive, there is
human fatigue caused by the interaction \cite{ie1}, and also boredom arise
when users evaluate a large number of phenotypes, 
many of which are not interesting or are very similar to each other.
% C-IEAs
Moreover, the performance of these systems effectively depends on the number of users
they are able to include; to reach more users,
IEC systems are some times developed as web applications depending on web visitors to help
with the search, using both anonymous and registered users. Some systems 
employ a collaborative technique, where several users participate in 
the evaluation, this method is called Collaborative-IEC (C-IEC)
\cite{picbreeder,seyama2016development,wagy2014collective}.
% Volunteers 
Having a volunteer system can lower the requirements of anyone participating in
the experiment thus increasing the {\em performance}, in terms of 
{\em human brain cycles}, of the whole system. But using a volunteer based 
system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
volunteer\'s lack of accountability, and the need to build trust between participants and project
owners. Other issues for project owners are also the difficulty of establishing 
the amount of time and resources a volunteer is willing to spend on the system, 
and how they decide if they participate or not \cite{JJ:2016}.

@@

For example, in a C-IEC application, fitness assignment depends on the
actions of a social network of users.  These actions are triggered when 
they tag, share, rate, store or even delete a phenotype. 
Then, the selection of parents could depend on the previous actions, leveraging information 
such as the fact that they both have the same tag, or where shared by
similar users.

Data available from the interaction is also used to increase the engegment of 
users in the system by applying  gamification techniques. Gamification is defined by 
Deterding et al. \cite{deterding2011game} as
\begin{quote}
  the use of game design elements in non-game contexts.
\end{quote}  
The gamification element employed in this case study has been a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level. % Say what we are doing in this case - JJ
A case study consisting in a web application for evolving artistic drawings
is presented, providing both conceptual and implementation details of the framework. 
In order to test the framework three versions of the application are compared to
measure the degree of relationship between users and their participation,
when appling a gamification technique.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution. Then, Section \ref{sec:framework} presents the framework of Human
Interactions in Collaborative Interactive Evolutionary Computation (C-IEC) which 
the main proposal of this work. An implementation using the ES-I framework 
is detailed in Section \ref{sec:implementation}.
The case study and results are presented in Sections \ref{sec:experiments} and  
\ref{sec:results}. Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

>>>>

% IEAs
Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
whose fitness evaluations are performed by humans through an interactive 
system \cite{eiben2015interactive}.
They are usually applied in problems where the fitness function is not known or simply
does not exist, and the result of optimization should fit a certain human need or desire. 
That is why their use-case includes the evolution of objects with subjective characteristics
such visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
or creativity \cite{yu2011cooks}.
% Human Based Computation
These IEC systems are an interesting venue of research, since they have demonstrated 
their ability for effectively 
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
as well as other artifacts in many other domains \cite{ie1}. In those cases when 
human interaction is responsible of other 
aspects of the evolutionary process, some authors classify these IEC methods 
as human-based evolutionary computation \cite{kosorukoff2001human} 
or as human-based computation \cite{quinn2011human}.
% Problems
However, the necessary intervention of humans, brings certain challenges 
to designers of IEC methods namely, human evaluations are slow and expensive, there is
human fatigue caused by the interaction \cite{ie1}, and also boredom arise
when users evaluate a large number of phenotypes, 
many of which are not interesting or are very similar to each other.
% C-IEAs
Moreover, the performance of these systems effectively depends on the number of users
they are able to include; to reach more users,
IEC systems are some times developed as web applications depending on web visitors to help
with the search, using both anonymous and registered users. Some systems 
employ a collaborative technique, where several users participate in 
the evaluation, this method is called Collaborative-IEC (C-IEC)
\cite{picbreeder,seyama2016development,wagy2014collective}.
% Volunteers 
Having a volunteer system can lower the requirements for anyone to participate in
the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
volunteer\'s lack of accountability, and the need to build trust between participants and project
owners. Other issues for project owners are also the difficulty of establishing 
the amount of time and resources a volunteer is willing to spend on the system, 
and how they decide if they participate or not \cite{JJ:2016}.

@@

For example, in a C-IEC application fitness assignment depends on the
actions of a social network of users.  These actions are triggered when 
they tag, share, rate, store or delete a phenotype. 
Then, the selection of parents could depend on the previous actions, leveraging information, 
such as the fact that they both have the same tag, or were shared by
similar users.

Data available from the interaction is also used to increase the engegment of 
users in the system by applying  gamification techniques. Gamification is defined by 
Deterding et al. \cite{deterding2011game} as
\begin{quote}
  the use of game design elements in non-game contexts.
\end{quote}  
The gamification element employed in this case study has been a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level. % Say what we are doing in this case - JJ
%%%
%%% Change if case studies change
%In this work, a case study consisting in a web application for evolving artistic drawings
%is presented, providing both conceptual and implementation details of the framework. 
%In order to test the framework three versions of the application are compared to
%measure the degree of relationship between users and their participation,
%when appling a gamification technique. 
In this work, three case studies are presented. First, the development of a web based C-IEC application 
for evolving artistic drawings is described, to give the reader details about 
the utilization of the proposed framework. Then, a rewarding mechanism is 
implemented in the same application, to study the changes in participation,
when appling a gamification technique. Later two other case studies are briefly presented
to show how the framework is applied in other domains.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution. Then, Section \ref{sec:framework} presents the human centered framework of
for C-IEC appliactions which the main proposal of this work. Next three case studies are presented in
sections

An implementation using the ES-I framework 
is detailed in Section \ref{sec:implementation}.
The case study and results are presented in Sections \ref{sec:experiments} and  
\ref{sec:results}. Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
